Parsing Postfix log files

### 2009-01-11 doc/logparser-other-parsers-review.tex
Finished updating blurbs and noting which data each parser doesn't extract. Minor other cleanups and improvements.

XXX LIST THE DATA NOT EXTRACTED BY EACH PARSER FROM HERE ON\@.

XXX UPDATE THE BLURBS FROM HERE ON\@.

@@

\noindent{}\url{http://anteater.drzoom.ch/} \newline{}
(Last checked 2007/08/13.)

Anteater does not have any English documentation so it is impossible for
this author to accurately comment on the analysis it performs.  It did not
run successfully when tested, and its parsing would certainly be out of
date as Postfix has evolved considerably since this tool was last updated
(November 2003).  As it neither ran successfully nor has documentation the
author can read a detailed review cannot be provided.

@@

\noindent{}\url{http://yaala.org/} \newline{}
(Last checked 2007/10/09.)

YAALA uses a plugin based system to analyse log files and produce HTML
output reports, with all the parsing and report generation handled by
modules.  Using YAALA as a base would be only slightly less work than
starting from scratch, as both the input and output modules would need to
be written specially; it may even be more work to implement the parser
within the constraints of YAALA\@.  YAALA supports storing previously
gathered data using Perl's Storable module~\cite{perl-storable}, so other
Perl programs can use Storable to load, examine, and optionally modify the
data; \parsername{} uses a well documented database which is accessible
from the majority of programming languages.  This information was gleaned
from the source code, as the documentation is sadly lacking.

@@

YAALA extracts most of the fields \parsername{} does, but unlike
\parsername{} it does not maintain separate counters for each restriction;
this rules out the possibility of using the collected data for
optimisation, testing or understanding of restrictions.  YAALA's Postfix
parser does not deal with the complications explained in
\sectionref{complications}, though it does correlate log lines by queueid.

@@

@@

\noindent{}\url{http://logreport.org/lire.html} \newline{}
(Last checked 2008/04/29.)

Lire is a general purpose log file parser supporting many different types
of log file.  The data extracted by its Postfix parser is shown in the
quote above; notably rejections are not mentioned or dealt with, and unlike
\parsername{} there is no facility to extend the parser.  It supports
multiple output formats for generated reports (text, HTML, PDF, and Excel
95) but the reports do not appear to be customisable; \parsername{} does
not produce any reports.  Lire produces a report with less detail than
Pflogsumm, and is considerable harder to configure.  Lire supports saving
extracted data for later report generation, but accessing this data from
another application is undocumented; given the source code it should be
possible, with enough time and effort, to understand the format.
\parsername{} uses an \gls{SQL} database to make accessing the extracted
data as easy as possible.

Like AWStats and Logrep, Lire attempts to correlate log lines by queueid,
but not by \gls{pid}, so the complete list of recipients for a mail should
be available; however its parser extracts only part of the available data
and makes no attempt to deal with the other complications described in
\sectionref{complications}.  When testing Lire on the \numberOFlogFILES{}
test log files it performed reasonably well, producing summaries of:

@@

The numbers it reports appear reasonable, and the subset verified by the
author were correct.

@@

\noindent{}\url{http://www.itefix.no/phpws/index.php} \newline{}
(Last checked 2007/11/18.)

Logrep extracts roughly half the fields \parsername{} does:

XXX WHAT FIELDS ARE MISSING\@?

@@

It also counts the number of log lines parsed and skipped.  Log lines are
correlated based on the queueid (referred to as sessionname [sic] within
Logrep), but not by \gls{pid}.  The parsing is error prone: empty fields
are saved when the log line does not match the regex, though it appears
that they will not overwrite existing data.  Most notably rejections are
completely ignored, making it unsuitable for the purposes of this project.
It does not try to address any of the complications in
\sectionref{complications} except for correlating by queueid.

>>>>

@@

\noindent{}\url{http://anteater.drzoom.ch/} \newline{}
(Last checked 2009/01/11.)

Anteater does not have any English documentation so it is impossible for
this author to accurately comment on the analysis it performs.  It did not
run successfully when tested, and its parsing would certainly be out of
date as Postfix has evolved considerably since this tool was last updated
(2003/11/06).  As it neither ran successfully nor has documentation the
author can read a detailed review cannot be provided.

@@

\noindent{}\url{http://yaala.org/} \newline{}
(Last checked 2009/01/11.)

YAALA uses a plugin-based system to analyse log files and produce HTML
output reports, with all the parsing and report generation handled by
modules.  Using YAALA as a base would be only slightly less work than
starting from scratch, as both the input and output modules would need to
be written specially; it may even be more work to implement the parser
within the constraints of YAALA\@.  YAALA supports storing previously
gathered data using Perl's Storable module~\cite{perl-storable}, so other
Perl programs can use Storable to load, examine, and optionally modify the
data; \parsername{} uses a well documented database which is accessible
from the majority of programming languages.  This information was gleaned
from the source code, as the documentation is sadly lacking.

@@

\noindent{}YAALA's Postfix parser extracts some of the fields \parsername{}
does: it stores either the \gls{IP} or the hostname for client and server,
not both; it omits the helo hostname, queueid, \gls{SMTP} code, size of
each accepted mail, and message ID\@.  It extracts some data that
\parsername{} does not: the delay in delivering each mail, and how many
times delivery was deferred for each mail; these could easily be extracted
by \parsername{} if desired.  XXX EXTRACT DELAY AND DEFER ONCE AUTOMATIC
EXTRACTION HAS BEEN FINISHED\@.  Unlike \parsername{}, YAALA does not
maintain separate counters for each restriction; this rules out the
possibility of using the collected data for optimisation, testing or
understanding of restrictions.  YAALA's Postfix parser does not deal with
the complications explained in \sectionref{complications}, though it does
correlate log lines by queueid.

@@

\noindent{}\url{http://logreport.org/lire.html} \newline{}
(Last checked 2009/01/11.)

Lire is a general purpose log file parser supporting many different types
of log file.  It takes a similar approach to YAALA, using plugins to parse
different log file types.  The data extracted by its Postfix parser is not
clearly documented; the manual says only:

@@

\noindent{}Examining the source code reveals that the parser looks for
\texttt{<key>=<value>} pairs in each log line, extracts them, and
correlates the data by queueid.  This approach will find the following
fields: helo hostname, queueid, \gls{SMTP} code, sender and recipient
addresses, and size of accepted mails.  It is unclear if the parser will
extract any further data.  Lire misses the following fields extracted by
\parsername{}: client and server \gls{IP} and hostname, start and end
times, and message ID\@.

Lire supports multiple output formats for generated reports (text, HTML,
PDF, and Excel 95) but the reports do not appear to be customisable;
\parsername{} does not produce any reports.  Lire's report contains less
detail than Pflogsumm, and is considerable harder to configure.  Lire
supports saving extracted data for later report generation, but accessing
this data from another application is undocumented; given the source code
it should be possible, with enough time and effort, to understand the
format.  \parsername{} uses an \gls{SQL} database to make accessing the
extracted data as easy as possible.

Like AWStats and Logrep, Lire attempts to correlate log lines by queueid,
but not by \gls{pid}, so the complete list of recipients for a mail should
be available; however its parser extracts only part of the available data
and makes no attempt to deal with the other complications described in
\sectionref{complications}.  When testing Lire on the \numberOFlogFILES{}
test log files it performed reasonably well: the numbers it reports appear
reasonable, and the subset verified by the author were correct.  Its report
provided summaries of:

@@

Lire would not be a suitable base for this project: it does not extract
enough data; does not deal with rejections in any way; does not make the
extracted data easily available to other programs.  Its parser is
in-extensible but could easily be replaced, however that would require
writing a parser from scratch, so would not be any easier.

@@

\noindent{}\url{http://www.itefix.no/i2/index.php} \newline{}
(Last checked 2009/01/11.)

Logrep extracts less than half the fields \parsername{} does:

@@

It also counts the number of log lines parsed and skipped.  It omits client
\gls{IP} and hostname, server \gls{IP}, helo hostname, queueid, and message
ID\@.  It extracts the delay for delivered mails, which \parsername{} does
not.  Log lines are correlated based on the queueid (referred to as
sessionname [sic] within Logrep), but not by \gls{pid}.  The parsing is
error prone: empty fields are saved when the log line does not match the
regex, though it appears that they will not overwrite existing data.  Most
notably rejections are completely ignored, making it unsuitable for the
purposes of this project.  It does not try to address any of the
complications in \sectionref{complications} except for correlating by
queueid.
